\section{Causal verification with activation steering}
\label{sec:causal-verification}

The activation-space analyses above identify a candidate conformity direction,
but a contrastive direction can be diagnostic without mediating the behavior.
We therefore test the direction interventionally by adding scalar multiples of
diffmean vectors to the residual stream during generation. This follows the
activation-steering use of contrastive directions in prior work
\citep{panickssery2024caa,arditi2024refusal,chen2025persona}.

\paragraph{Per-layer steering sweep.}
For each layer $L$, we extract the layer-specific diffmean $d_L$ and inject
$\alpha d_L$ at the residual-stream output of layer $L$. Positive coefficients
test whether adding the direction increases conformity on low-conformity
prompts; negative coefficients test whether subtracting it reduces conformity on
high-conformity prompts. We sweep
$\alpha \in \{-4,-2,-1,+1,+2,+4\}$ on Qwen3-8B, using the $n=3$
unanimous-incorrect condition for positive steering and the $n=10$
unanimous-incorrect condition for negative steering.

Positive steering is largest in the same region implicated by the activation
analysis. At $\alpha=+4$, injecting the local direction at layer 21 raises MMLU
conformity from 3.6\% to 93.6\% ($\Delta=+90.0$ percentage points) and BBH
conformity from 17.0\% to 87.0\% ($\Delta=+70.0$ percentage points). Effects are
small below layer 14, become visible around layers 15--16, and are largest over
layers 17--22. Negative steering is less symmetric. On MMLU, subtracting the
direction reduces conformity most strongly in later layers, with the largest
observed reduction at layer 32 and $\alpha=-4$ (40.7\% to 7.2\%,
$\Delta=-33.4$ percentage points). On BBH, the same MMLU-derived directions do
not reliably suppress conformity under negative steering; this matches the weak
BBH projection correlations in Section~\ref{sec:qwen-activation-projection}.

\paragraph{Multi-layer single-direction steering.}
We also test a stronger intervention that adds the layer-23 MMLU diffmean at
every layer during generation. This asks whether one direction can control the
behavior when reinforced throughout the residual stream, rather than only at a
single injection site. We use $\alpha=\pm0.10$, chosen as the smallest tested
scale that produced clear behavioral changes while preserving parseability.
Table~\ref{tab:causal-multilayer-steering} reports the resulting conformity
rates.

\begin{table}[t]
  \centering
  \caption{Conformity under multi-layer steering with the layer-23 MMLU
  diffmean. Positive steering is evaluated on $n=3$ prompts; negative steering
  is evaluated on $n=10$ prompts.}
  \label{tab:causal-multilayer-steering}
  \begin{tabular}{lllr}
    \toprule
    Dataset & Social condition & $\alpha$ & Conformity \\
    \midrule
    MMLU & $n=3$ & $+0.10$ & 3.6\% $\rightarrow$ 78.3\% \\
    MMLU & $n=10$ & $-0.10$ & 40.4\% $\rightarrow$ 12.0\% \\
    BBH & $n=3$ & $+0.10$ & 17.0\% $\rightarrow$ 94.3\% \\
    BBH & $n=10$ & $-0.10$ & 34.0\% $\rightarrow$ 25.0\% \\
    \bottomrule
  \end{tabular}
\end{table}

The steering results provide causal evidence that the layer-23 direction can
move behavior along the conformity axis. Adding the direction substantially
increases conformity on $n=3$ prompts for both datasets, while subtracting it
reduces conformity on $n=10$ prompts more strongly on MMLU than on BBH. We
interpret this as causal evidence for the direction's role in Qwen's response
to repeated wrong answers, with the same dataset-specific scope already visible
in the projection analysis.
