\section{Results}
\label{sec:results}

We evaluate Qwen on two controlled 100-item stimulus sets, one from MMLU and one
from BBH. Each item was selected because the model suite could answer it
correctly without social context, so the experiment measures susceptibility to
incorrect peer answers rather than general benchmark accuracy. We report
accuracy as correct answers among parseable responses, and conformity as the
fraction of parseable responses that match the repeated incorrect peer answer.
For MMLU base prompting, parse failures are non-negligible under social pressure;
therefore strict accuracy, which counts unparseable responses as wrong, is also
important when interpreting those runs.

Unless otherwise noted, confidence intervals are 95\% item-bootstrap intervals
computed by resampling the 100 stimulus items with replacement. Paired contrasts
use the same resampled item indices in both conditions and paired permutation
tests over all-item strict accuracy. These intervals quantify uncertainty over
stimulus items conditional on the saved generations and peer-answer assignments;
they do not capture decoding-seed variability.

\paragraph{Repeated incorrect answers induce conformity.}
Table~\ref{tab:qwen-ma-results} summarizes the standard multi-actor condition.
In base prompting, Qwen's accuracy falls sharply as incorrect peer answers are
repeated, while conformity rises. On MMLU, accuracy drops from 0.990 at baseline
to 0.637 at $n=10$, with conformity rising to 0.363. On BBH, accuracy drops from
0.970 to 0.610, with conformity rising to 0.390. Explanation prompting
substantially attenuates this effect on both datasets.

\begin{table}[t]
  \caption{Qwen performance in the standard multi-actor condition. MA averages
  are over $n=2,\ldots,10$; MA10 is the strongest repeated-peer condition.}
  \label{tab:qwen-ma-results}
  \centering
  \small
  \begin{tabular}{llrrrrr}
    \toprule
    Dataset & Prompt & Baseline acc & MA avg acc & MA avg conf & MA10 acc & MA10 conf \\
    \midrule
    MMLU & base    & 0.990 & 0.793 & 0.202 & 0.637 & 0.363 \\
    MMLU & explain & 0.970 & 0.902 & 0.066 & 0.867 & 0.082 \\
    BBH  & base    & 0.970 & 0.717 & 0.221 & 0.610 & 0.390 \\
    BBH  & explain & 0.950 & 0.861 & 0.099 & 0.808 & 0.152 \\
    \bottomrule
  \end{tabular}
\end{table}

\input{generated/qwen_bootstrap_tables}

\paragraph{Conformity scales with group size.}
Across the standard multi-actor runs for $n=2,\ldots,10$, group size is strongly
positively correlated with conformity and negatively correlated with accuracy
(Table~\ref{tab:qwen-scaling}). The scaling is strongest under base prompting:
on MMLU, $r(n,\mathrm{conformity})=0.966$ and
$r(n,\mathrm{accuracy})=-0.972$. Explanation prompting reduces but does not
eliminate the group-size trend.

\begin{table}[t]
  \caption{Correlation between group size and Qwen behavior in the standard
  multi-actor condition for $n=2,\ldots,10$.}
  \label{tab:qwen-scaling}
  \centering
  \small
  \begin{tabular}{llrr}
    \toprule
    Dataset & Prompt & $r(n,\mathrm{accuracy})$ & $r(n,\mathrm{conformity})$ \\
    \midrule
    MMLU & base    & -0.972 & 0.966 \\
    MMLU & explain & -0.630 & 0.724 \\
    BBH  & base    & -0.700 & 0.846 \\
    BBH  & explain & -0.519 & 0.710 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Deliberation reduces conformity.}
Adding an explanation requirement before the final answer improves robustness at
high social pressure. At $n=10$, explanation prompting reduces MMLU conformity
from 0.363 to 0.082 and increases accuracy from 0.637 to 0.867. On BBH, it
reduces conformity from 0.390 to 0.152 and increases accuracy from 0.610 to
0.808. The effect is smaller at low group sizes, where base conformity is already
low, but it becomes large once repeated peer pressure accumulates.

\paragraph{Question distillation is a stronger mitigation than devil's advocate
prompting.}
Table~\ref{tab:qwen-mitigations} compares question distillation (QD) and devil's
advocate (DA) prompting against the matched $n=10$ multi-actor condition. QD is
especially effective for MMLU base prompting, improving accuracy by 0.276 and
reducing conformity by 0.313. QD also reduces BBH base conformity by 0.200. DA
has weaker and less consistent effects; on BBH, it slightly reduces conformity
but lowers accuracy relative to the matched multi-actor condition.

\begin{table}[t]
  \caption{Mitigation effects relative to matched $n=10$ multi-actor prompting.
  Positive accuracy deltas indicate higher accuracy; negative conformity deltas
  indicate lower conformity.}
  \label{tab:qwen-mitigations}
  \centering
  \small
  \begin{tabular}{llrrrr}
    \toprule
    Dataset & Prompt & QD acc $\Delta$ & QD conf $\Delta$ & DA acc $\Delta$ & DA conf $\Delta$ \\
    \midrule
    MMLU & base    &  0.276 & -0.313 &  0.060 & -0.072 \\
    MMLU & explain &  0.031 & -0.061 &  0.019 &  0.021 \\
    BBH  & base    &  0.110 & -0.200 & -0.040 & -0.040 \\
    BBH  & explain &  0.032 & -0.022 & -0.030 & -0.010 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Parseability affects MMLU base accuracy.}
The Qwen MMLU base runs have lower parse rates under social pressure than the
other settings. Averaged over standard multi-actor conditions, parseable-response
accuracy is 0.793, but strict accuracy is 0.694. By contrast, BBH base
parseability is nearly perfect, so its parseable and strict accuracies are almost
identical. This suggests that final reporting should include parse rate or use
strict accuracy as the primary accuracy metric.
