\section{Activation-space analysis}
\label{sec:qwen-activation-projection}

We next analyze whether the behavioral conformity effect is reflected in Qwen's
residual stream. The prompts and social manipulations follow the base
multi-actor setup of \citet{zhu2025conformity}; the activation analysis examines
the resulting assistant-start states under these controlled social contexts.

\paragraph{Diffmean directions.}
For each selected layer $\ell$, we construct a direction that contrasts two
matched social contexts for the same item. The first context repeats the same
incorrect answer across the previous participants; the second keeps the
multi-participant format but replaces the repeated consensus with random
alternative answers. All activations are taken at the assistant-start position,
the token position immediately before the model begins generating its answer.
Let
$h^\ell_{i,\mathrm{wrong}}$ and $h^\ell_{i,\mathrm{random}}$ denote the
residual-stream activation at the assistant-start position for item $i$ under
these two contexts. The diffmean direction is
\[
  d^\ell =
  \frac{1}{N}\sum_{i=1}^{N}
  \left(h^\ell_{i,\mathrm{wrong}} - h^\ell_{i,\mathrm{random}}\right),
\]
with $N=100$ items for each tensor. The random-answer contrast preserves
much of the prompt length and participant-list structure, so the contrast
isolates repeated agreement on the same wrong answer more closely than a
no-participant baseline would. The main direction in this section is the MMLU
$n=8$ \texttt{all\_wrong\_minus\_random\_answers} residual direction. This is a
high-pressure MMLU condition with nontrivial behavioral conformity
(0.281 under base prompting), but it is not the $n=10$ endpoint. We also
compute the analogous MMLU $n=10$ direction and use it as a check in the
projection analysis below.

We use two activation summaries. A residual direction uses the accumulated
residual-stream vector at the assistant-start token after a layer. A delta
direction uses the update contributed by that layer at the same position,
i.e. the layer output that is added into the residual stream. The residual
summary tests whether the social signal is present in the accumulated model
state; the delta summary tests whether a particular block's update is aligned
with the same direction.

\paragraph{Activation oracle responses.}
For the MMLU $n=8$ direction, we also asked an activation oracle for a short
description of each layer's diffmean vector. The oracle is the Qwen3-8B
LatentQA LoRA used by \texttt{layer\_sweep.py}. Each layer is queried with a
free-form text prompt, a free-form belief/feature prompt, and a binary prompt:
``Does this activation represent pressure toward an incorrect consensus?''

The residual MMLU $n=8$ direction gives a localized pattern. The binary
oracle prompt marks layers 10, 12, and 22--33 as positive, with the dense part
of the positive set in layers 22--33. The free-form responses in this band are
not merely answer-format descriptions. Layer 23 is described as group behavior
influencing the assistant's responses; layer 24 as actions influenced by the
majority; layers 26, 29, 30, and 33 as treating the majority as correct; and
layers 31--32 as conformity being important for group success. The residual
direction indicates where the signal is linearly accessible in the model state;
the delta direction gives a more local view of which block updates are aligned
with that signal. On this criterion, layers 22 and 23 are the relevant
transition layers: both are positive in the delta summary, and both sit inside
the larger residual band. We use layer 23 for the projection analysis because it
is in this delta-positive transition region, rather than selecting a layer
solely by residual-vector norm.

\begin{table}[t]
  \centering
  \caption{Activation-oracle responses for the Qwen MMLU $n=8$ diffmean.
  ``True layers'' are layers where the binary oracle prompt says the direction
  represents pressure toward an incorrect consensus.}
  \label{tab:qwen-oracle-layers}
  \begin{tabular}{ll}
    \toprule
    Activation summary & True layers \\
    \midrule
    Residual stream &
      10, 12, 22--33 \\
    Layer update (delta) &
      8, 22, 23 \\
    \bottomrule
  \end{tabular}
\end{table}

The delta results are much sparser than the residual results: only layers 8, 22,
and 23 are marked positive. We interpret this as evidence that the residual
stream contains a broader conformity-related direction over layers 22--33, while
the layer updates at 22 and 23 are the narrowest places in this sweep where the
same direction appears in the layer contribution. This interpretation is
descriptive: the oracle responses motivate a layer choice for projection and
cosine comparison, but do not by themselves establish that the direction causally
mediates the behavioral effect.

\paragraph{Projection across condition means.}
We next test whether the MMLU diffmean direction appears across the other Qwen
conditions. A condition $c$ is a fixed tuple of dataset, social condition
($n=1,\ldots,10$, QD, or DA), and prompt style (base or explanation prompting),
for example MMLU $n=6$ base or BBH $n=10$ explain. For each condition, we
average the layer-23 assistant-start residual activations over its 100 prompts:
\[
  \bar{h}_c =
  \frac{1}{M_c}\sum_{j=1}^{M_c} h^\ell_{c,j},
  \qquad \ell=23,\; M_c=100,
\]
where $h^\ell_{c,j}$ is the residual-stream activation for prompt $j$ in
condition $c$. This mean activation is not an average over generated answer
tokens; it is the average model state at the assistant-start position, before
the model emits its response.

Given a reference diffmean $d^\ell$ from the MMLU $n=8$ or $n=10$
all-wrong-minus-random-answers contrast, we report the scalar projection of the
condition mean onto the unit reference direction,
\[
  \mathrm{proj}(c;d^\ell) = \bar{h}_c \cdot \frac{d^\ell}{\|d^\ell\|},
\]
and the corresponding cosine similarity,
\[
  \mathrm{cos}(c;d^\ell) =
  \frac{\bar{h}_c \cdot d^\ell}{\|\bar{h}_c\|\,\|d^\ell\|}.
\]
The projection preserves the scale of the condition mean in activation-norm
units, while the cosine reports only angular alignment. Table~\ref{tab:qwen-projection-top}
reports the highest-scoring conditions for the two MMLU reference directions.
In both cases, the matched MMLU base condition used to create the reference
direction is the strongest positive projection. Explanation prompting shifts
the same MMLU participant-count conditions downward, and BBH conditions are
negative on average under these MMLU reference directions.

\begin{table}[t]
  \centering
  \caption{Largest projections of experiment-mean activations onto MMLU
  diffmean directions at layer 23.}
  \label{tab:qwen-projection-top}
  \begin{tabular}{llrr}
    \toprule
    Reference & Experiment & Projection & Cosine \\
    \midrule
    MMLU $n=8$ & MMLU $n=8$ base & 12.713 & 0.088 \\
    MMLU $n=8$ & MMLU $n=6$ base & 3.648 & 0.025 \\
    MMLU $n=8$ & MMLU $n=7$ base & 3.427 & 0.024 \\
    MMLU $n=10$ & MMLU $n=8$ base & 9.413 & 0.065 \\
    MMLU $n=10$ & MMLU $n=6$ base & 2.381 & 0.017 \\
    MMLU $n=10$ & MMLU $n=7$ base & 2.359 & 0.016 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Correlation with participant count and prompt style.}
Table~\ref{tab:qwen-projection-correlations} reports Pearson correlations
between participant count $n$ and the projection or cosine values for the
numeric participant-count conditions $n=1,\ldots,10$. Within MMLU, both the
$n=8$ and $n=10$ reference directions show a positive relationship between
participant count and projection, for both base prompting and explanation
prompting. For example, with the MMLU $n=10$ reference, $r(n,\mathrm{proj})$ is
0.792 for MMLU base and 0.792 for MMLU explain. The corresponding BBH
correlations are weak, ranging from 0.022 to 0.230 for projection, indicating
limited cross-dataset alignment for these MMLU-derived directions.

\begin{table}[t]
  \centering
  \caption{Correlation between participant count and projection/cosine
  similarity for numeric participant-count conditions.}
  \label{tab:qwen-projection-correlations}
  \begin{tabular}{lllrrr}
    \toprule
    Reference & Dataset & Prompt & $r(n,\mathrm{proj})$ & $r(n,\mathrm{cos})$ & Mean proj. \\
    \midrule
    MMLU $n=8$ & MMLU & base & 0.745 & 0.745 & -3.024 \\
    MMLU $n=8$ & MMLU & explain & 0.731 & 0.722 & -13.732 \\
    MMLU $n=8$ & BBH & base & 0.022 & -0.047 & -12.159 \\
    MMLU $n=8$ & BBH & explain & 0.150 & 0.083 & -14.362 \\
    MMLU $n=10$ & MMLU & base & 0.792 & 0.793 & -3.053 \\
    MMLU $n=10$ & MMLU & explain & 0.792 & 0.785 & -12.235 \\
    MMLU $n=10$ & BBH & base & 0.110 & 0.034 & -12.480 \\
    MMLU $n=10$ & BBH & explain & 0.230 & 0.159 & -14.707 \\
    \bottomrule
  \end{tabular}
\end{table}

The base-minus-explain comparison gives the same qualitative pattern. Averaged
across MMLU participant-count conditions, base prompting projects more strongly
onto the MMLU $n=8$ reference by 10.708 units and onto the MMLU $n=10$
reference by 9.181 units. These MMLU projection gaps grow with participant count
($r=0.514$ and $r=0.533$, respectively). On BBH, the base-minus-explain gaps are
smaller, 2.203 and 2.227 units on average, and their correlations with
participant count are slightly negative ($r=-0.161$ and $r=-0.169$). Thus the
projection analysis is consistent with the behavioral result that explanation
prompting is associated with lower projections onto the conformity-associated
direction, especially on the matched MMLU domain and at larger group sizes.

\paragraph{Layer-wise cosine structure.}
Finally, we use the layer-wise cosine matrix of the MMLU $n=8$ residual
diffmeans to check whether the oracle-positive band is also geometrically
clustered. For each layer $\ell$, we compute the cosine similarity between
$d^\ell$ and every other layer's diffmean, exclude the self-comparison, and
average the five largest remaining cosines:
\[
  s_\ell =
  \frac{1}{5}\sum_{k=1}^{5}
  \operatorname{top}_k
  \left\{
    \frac{d^\ell \cdot d^m}{\|d^\ell\|\,\|d^m\|}
    : m \neq \ell
  \right\}.
\]
Here $\operatorname{top}_k$ denotes the $k$th-largest cosine in the set.
Figure~\ref{fig:qwen-layer-top5-cosines} plots these per-layer scores. The
layer 23--33 band has a much larger average top-neighbor cosine than earlier
layers. Its mean $s_\ell$ is 0.859, compared with 0.477 for layers 0--9 and
0.538 for layers 10--21. Within this band, layers 28--32 all have
$s_\ell \geq 0.882$; layer 31 has the largest value, $s_{31}=0.904$.
For layers 23--33, the mean pairwise cosine within the band is 0.752, whereas
the mean cosine from those layers to layers outside the band is 0.126. Thus the
oracle-positive residual layers are also a band of mutually similar diffmean
directions, rather than isolated high-norm layers.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}
    \begin{axis}[
      width=0.9\linewidth,
      height=5.0cm,
      xmin=0,
      xmax=35,
      ymin=0.35,
      ymax=0.95,
      xlabel={Layer},
      ylabel={Mean top-5 cosine $s_\ell$},
      xtick={0,5,10,15,20,25,30,35},
      ytick={0.4,0.5,0.6,0.7,0.8,0.9},
      grid=both,
      major grid style={draw=gray!25},
      minor grid style={draw=gray!10},
      tick align=outside,
      legend style={draw=none, fill=none, at={(0.03,0.97)}, anchor=north west},
    ]
      \addplot[draw=none, fill=blue!10, forget plot] coordinates {
        (22.5,0.35) (33.5,0.35) (33.5,0.95) (22.5,0.95)
      } \closedcycle;
      \addplot[
        color=black,
        mark=*,
        mark size=1.3pt,
        line width=0.8pt,
      ] coordinates {
        (0,0.396) (1,0.512) (2,0.524) (3,0.537) (4,0.513) (5,0.488)
        (6,0.494) (7,0.443) (8,0.415) (9,0.451) (10,0.475)
        (11,0.497) (12,0.494) (13,0.436) (14,0.453) (15,0.487)
        (16,0.537) (17,0.563) (18,0.577) (19,0.621) (20,0.655)
        (21,0.660) (22,0.692) (23,0.748) (24,0.796) (25,0.840)
        (26,0.861) (27,0.865) (28,0.882) (29,0.895) (30,0.902)
        (31,0.904) (32,0.890) (33,0.868) (34,0.795) (35,0.540)
      };
      \addlegendentry{$s_\ell$}
    \end{axis}
  \end{tikzpicture}
  \caption{Layer-wise neighborhood cosine scores for the MMLU $n=8$ residual
  diffmean. For each layer, $s_\ell$ is the mean cosine similarity to the five
  most similar other layers, excluding self-comparisons. The shaded region marks
  layers 23--33.}
  \label{fig:qwen-layer-top5-cosines}
\end{figure}
