\documentclass{article}

\usepackage{neurips_2026}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

\title{Mechanistic Interpretability of Conformity in LLMs}

\author{Anonymous Authors}

\begin{document}

\maketitle

\begin{abstract}
LLMs increasingly operate in multi-agent and conversational settings, and recent
behavioral work shows that they can conform to a unanimous incorrect majority,
mirroring Asch's classical findings in humans. While mechanistic
interpretability has been used to characterize adjacent social-compliance
phenomena such as sycophancy, the internal representations that support
conformity remain less understood. We address this gap by replicating and
extending Asch-style conformity experiments in Qwen3-8B on MMLU and BIG-Bench
Hard items that the model answers correctly without social context. We find that
repeated incorrect peer answers reliably pull the model away from otherwise
available answers, that this effect grows with the number of agreeing
participants, and that explanation prompting and question distillation reduce
conformity. We then apply difference-of-means (DiffMean) analysis to residual
stream activations, comparing all-wrong peer contexts with random-peer-answer
controls, and use activation oracles, projection analysis, and layer-wise cosine
structure to localize a coherent representation of incorrect-consensus pressure
in a late-middle layer band. Finally, activation steering with this direction
both induces conformity in low-pressure prompts and suppresses it in
high-pressure prompts, suggesting that peer-majority conformity is supported by
a detectable and steerable residual-stream feature.
\end{abstract}

\section{Introduction}

Large language models are routinely placed in collaborative settings: they
summarize previous answers, review other agents' outputs, and participate in
multi-agent workflows. A model may know the answer in isolation, but abandon it
when an apparent group repeatedly states a different answer.

Prior work adapts classic social-psychology paradigms to language models
\citep{asch1951effects,zhu2025conformity} and finds conformity across knowledge
domains, with sensitivity to group size, unanimity, difficulty, source
characteristics, and uncertainty
\citep{bellina2026conformity,zhong2025disentangling}. These behavioral results
raise a mechanistic question: when a model conforms, can the social-pressure
manipulation be localized in internal representations?

We run a small but controlled replication and extension on Qwen3-8B. The
stimuli are 100 MMLU and 100 BBH items that Qwen answers correctly without
social context. This isolates cases where an incorrect social signal can pull
the model away from a reachable answer. We measure accuracy, parseability, and
conformity to the repeated wrong peer answer under group sizes
$n=1,\ldots,10$, with and without an explanation requirement, plus question
distillation and devil's-advocate mitigations from \citet{zhu2025conformity}.

We then examine Qwen's internal activations. Following work on contrastive
activation directions \citep{panickssery2024caa,haon2025vla}, we construct
difference-of-means directions between repeated-wrong and random-answer peer
contexts at the assistant-start token, then test these directions with
projection, layer-wise cosine analysis, and activation steering.

\section{Related work}
\label{sec:related-work}

\paragraph{Conformity in large language models.}
Asch's classical conformity experiments established that a unanimous incorrect
majority can drive substantial conformity in human subjects, rising under task
uncertainty and collapsing in the presence of even a single dissenter
\citep{asch1956studies}. Recent work has begun translating this work to LLMs to
uncover whether they display similar conforming behavior. \citet{zhu2025conformity}
adapt Asch's experiments over eight open-source models and six question-answer
datasets to find that the majority of LLMs conform to a unanimous incorrect
majority and that LLM conformity scales inversely with model confidence. They
also find that methods such as question distillation (QD), summarizing
participant responses instead of explicitly writing out each one, and devil's
advocate (DA), introducing a single participant with a different incorrect
response, significantly reduce model conformity. \citet{weng2025benchform}
formalize the phenomenon over multi-round interaction with BENCHFORM, a
benchmark derived from the BIG-Bench Hard dataset with five interaction
protocols whose metrics, such as conformity rate and independence rate, we
adopt.

\paragraph{Sycophancy and behavior mitigation.}
Conformity sits within a broader family of social compliance behaviors that
includes sycophancy, originally surfaced as an inverse-scaling effect by
\citet{perez2023sycophancy} and traced by \citet{sharma2023sycophancy} to RLHF
preference data that rewards agreement with users. Behavior mitigations of
conformity and sycophancy have been proposed via prompting techniques to improve
reasoning and reflection \citep{weng2025benchform,sharma2023sycophancy}, yet
these approaches address the behavioral symptoms of social compliance, without
addressing the internal representations that produce them.

\paragraph{Linear representations and probing.}
A growing body of evidence supports the linear representation hypothesis:
high-level concepts in language models correspond to recoverable directions in
model activations \citep{park2024linear}. Linear directions have been identified
for truthfulness \citep{marks2024geometry,burns2023discovering}, honesty
\citep{li2023inference}, and a wide range of other behavioral traits via
interpretability methods. Linear probes have a long history as diagnostic tools
for what intermediate layers encode \citep{alain2017understanding}, though
\citet{belinkov2022probing} cautions that probes can succeed even when the
underlying property is not represented in the model, motivating the use of
low-capacity probes.

\paragraph{Steering vectors and DiffMean.}
\citet{turner2024activation} showed that simple contrasting natural-language
prompt pairs suffice to compute steering vectors, directions in the residual
stream that shift model behavior. \citet{panickssery2024caa} generalized this
into contrastive activation addition (CAA), the difference-of-means (DiffMean)
method which we adopt: residual-stream activations are averaged across positive
and negative behavior examples, and the resulting difference vector both reads
and steers the target model behavior. \citet{marks2024geometry} further showed
that DiffMean directions are more causally implicated than logistic regression
probes, providing warrant for our approach. Anthropic's persona vectors paper
\citep{chen2025persona} automates DiffMean for a number of character traits,
including sycophancy, to enable monitoring and steering to mitigate unwanted
behaviors.

\paragraph{Activation oracles and patching.}
More recently, activation oracles \citep{karvonen2025activation} have been used
to train LLMs to answer arbitrary natural-language questions about activations
from a target model. We use activation oracles to uncover where in the model
conforming behaviors are introduced, later using this layer-specific information
to calculate a DiffMean direction that encodes conformity. A direction's causal
role is established through intervention. Activation patching, replacing
activations from a clean run with those from a corrupted run with an added
direction, tests which components modify a behavior \citep{meng2022locating}.
Patching has also been used to reverse-engineer full circuits in controlled
tasks \citep{wang2023interpretability}.

\paragraph{Interpretability of LLM sycophancy.}
Recent mechanistic work on sycophancy shows that distinct sycophantic
sub-behaviors live along separable linear directions and admit independent
steering \citep{vennemeyer2025sycophancy}, and that user opinions override
learned model knowledge in mid-to-late layers on Qwen and Llama
\citep{li2025truthoverridden}. These studies, however, leave open whether the
conformity behavior shares a mechanism with sycophancy or constitutes a distinct
circuit. We close this gap by locating a conformity-specific direction in
Qwen3-8B, and testing its causal sufficiency via steering and activation
patching.

\input{sections/qwen_results_section.tex}

\input{sections/qwen_activation_projection_section.tex}

\input{sections/causal_verification.tex}

\section{Discussion}

The behavioral results reproduce the central conformity pattern of
\citet{zhu2025conformity}: repeated incorrect peer answers can pull Qwen away
from answers it can otherwise produce. Explanation prompting sharply reduces
conformity at large group sizes, and question distillation helps most clearly
for MMLU base prompting.

The activation results give a more tentative mechanistic picture. The
all-wrong-minus-random contrast controls for much of the generic effect of
adding peer-answer text. The activation oracle marks residual-stream layers 10,
12, and 22--33 as pressure toward an incorrect consensus, while the layer-update
results point to layers 22--23 as contributing the conformity direction.
Projections onto the layer-23 direction increase with participant count within
MMLU, explanation prompting shifts condition means downward, and layers 23--33
share a high-cosine MMLU diffmean direction. Steering gives an intervention
check: adding the layer-23 direction increases conformity on $n=3$ prompts, and
subtracting it reduces conformity on $n=10$ prompts, with stronger suppression
on MMLU than BBH.

\section{Limitations}
\label{sec:limitations}

The experiments use one open model, two selected 100-item stimulus sets, and one
stochastic generation per condition after retries. Confidence intervals quantify
item-level uncertainty only; they do not capture prompt paraphrase sensitivity,
seed-to-seed variability, model families, or arbitrary benchmark samples. The
selected-item design isolates social susceptibility but overstates baseline
accuracy relative to ordinary benchmark evaluation.

The activation oracle is itself a model component, and its labels are evidence
to inspect rather than ground truth. The projection and cosine analyses use
assistant-start diffmeans, so they can miss token-local mechanisms or more
distributed representations. Steering shows that the direction can influence
behavior under these interventions, but does not identify a complete mechanism.

\section{Reproducibility and ethics}
\label{sec:reproducibility}

The release repository contains the full code used for the experiments,
including generation scripts, activation-caching and DiffMean scripts,
activation-oracle scripts, activation-steering scripts, raw Qwen result files,
activation artifacts, and generated CSV/JSON summaries used for the behavioral
and activation figures. Section~\ref{sec:causal-verification} reports the
activation-steering protocol and summary results. The experiments use public
benchmarks and open model weights; no human subjects or crowdsourced workers are
involved.

All reported experiments were run on a node with two NVIDIA H100 GPUs. Behavioral
inference takes approximately 1 hour, activation patching takes approximately
2.5 hours, and activation caching, DiffMean construction, and activation-oracle
queries together take approximately 1 hour. All remaining analyses and figure
generation runs take under 30 minutes each.

The upstream assets are credited and used under their stated terms: Qwen/Qwen3-8B
is released under Apache-2.0, MMLU from \citet{hendrycks2021mmlu} is released
under the MIT License, and BIG-Bench Hard from \citet{suzgun2023bbh} is released
under the MIT License. The activation-oracle adapter
\texttt{adamkarvonen/checkpoints\_latentqa\_cls\_past\_lens\_addition\_Qwen3-8B}
is identified by Hugging Face repository ID and cited through
\citet{karvonen2025activation}; we do not redistribute Qwen weights or the
adapter weights, and users retrieve them from Hugging Face under the upstream
model-card terms. The released experiment code and derived paper artifacts are
distributed under the MIT License, subject to those upstream asset licenses.

The main ethical concern is deployment rather than data collection. The results
show that multi-agent or socially framed prompts can amplify a repeated wrong
answer, which is relevant to misinformation, biased consensus formation, and
automated decision pipelines that aggregate model outputs. The same mitigations
could be useful defensively, but exposing conformity measurements may also help
attackers design more persuasive prompts. We therefore frame the work as an
analysis of a failure mode and report limitations on generalization and control.

{\small
\begin{thebibliography}{99}

\bibitem[Alain and Bengio(2017)]{alain2017understanding}
Guillaume Alain and Yoshua Bengio.
\newblock Understanding intermediate layers using linear classifier probes.
\newblock In \emph{International Conference on Learning Representations
Workshop}, 2017.

\bibitem[Arditi et~al.(2024)]{arditi2024refusal}
Andy Arditi, Oscar Balcells Obeso, Aaquib Syed, Daniel Paleka, Nina Rimsky,
Wes Gurnee, and Neel Nanda.
\newblock Refusal in language models is mediated by a single direction.
\newblock In \emph{Advances in Neural Information Processing Systems}, 2024.

\bibitem[Asch(1951)]{asch1951effects}
Solomon E. Asch.
\newblock Effects of group pressure upon the modification and distortion of
judgments.
\newblock In H. Guetzkow, editor, \emph{Groups, Leadership and Men}, 1951.

\bibitem[Asch(1956)]{asch1956studies}
Solomon E. Asch.
\newblock Studies of independence and conformity: I. A minority of one against a
unanimous majority.
\newblock \emph{Psychological Monographs: General and Applied}, 70(9):1--70,
1956.

\bibitem[Bellina et~al.(2026)]{bellina2026conformity}
Alessandro Bellina, Giordano De Marzo, and David Garcia.
\newblock Conformity and social impact on AI agents.
\newblock \emph{arXiv preprint arXiv:2601.05384}, 2026.

\bibitem[Belinkov(2022)]{belinkov2022probing}
Yonatan Belinkov.
\newblock Probing classifiers: Promises, shortcomings, and advances.
\newblock \emph{Computational Linguistics}, 48(1):207--219, 2022.

\bibitem[Burns et~al.(2023)]{burns2023discovering}
Collin Burns, Haotian Ye, Dan Klein, and Jacob Steinhardt.
\newblock Discovering latent knowledge in language models without supervision.
\newblock In \emph{International Conference on Learning Representations}, 2023.

\bibitem[Chen et~al.(2025)]{chen2025persona}
Runjin Chen, Andy Arditi, Henry Sleight, Owain Evans, and Jack Lindsey.
\newblock Persona vectors: Monitoring and controlling character traits in
language models.
\newblock \emph{arXiv preprint arXiv:2507.21509}, 2025.

\bibitem[H\"aon et~al.(2025)]{haon2025vla}
Bear H\"aon, Kaylene Stocking, Ian Chuang, and Claire Tomlin.
\newblock Mechanistic interpretability for steering vision-language-action
models.
\newblock \emph{arXiv preprint arXiv:2509.00328}, 2025.

\bibitem[Hendrycks et~al.(2021)]{hendrycks2021mmlu}
Dan Hendrycks, Collin Burns, Steven Basart, Andy Zou, Mantas Mazeika, Dawn
Song, and Jacob Steinhardt.
\newblock Measuring massive multitask language understanding.
\newblock In \emph{International Conference on Learning Representations}, 2021.

\bibitem[Karvonen et~al.(2025)]{karvonen2025activation}
Adam Karvonen, James Chua, Cl\'ement Dumas, Kit Fraser-Taliente, Subhash
Kantamneni, Julian Minder, Euan Ong, Arnab Sen Sharma, Daniel Wen, Owain Evans,
and Samuel Marks.
\newblock Activation oracles: Training and evaluating LLMs as general-purpose
activation explainers.
\newblock \emph{arXiv preprint arXiv:2512.15674}, 2025.

\bibitem[Li et~al.(2023)]{li2023inference}
Kenneth Li, Oam Patel, Fernanda Vi\'egas, Hanspeter Pfister, and Martin
Wattenberg.
\newblock Inference-time intervention: Eliciting truthful answers from a
language model.
\newblock In \emph{Advances in Neural Information Processing Systems}, 2023.

\bibitem[Li et~al.(2025)]{li2025truthoverridden}
Jin Li, Keyu Wang, Shu Yang, Zhuoran Zhang, and Di Wang.
\newblock When truth is overridden: Uncovering the internal origins of
sycophancy in large language models.
\newblock \emph{arXiv preprint arXiv:2508.02087}, 2025.

\bibitem[Marks and Tegmark(2024)]{marks2024geometry}
Samuel Marks and Max Tegmark.
\newblock The geometry of truth: Emergent linear structure in large language
model representations of true/false datasets.
\newblock \emph{arXiv preprint arXiv:2310.06824}, 2024.

\bibitem[Meng et~al.(2022)]{meng2022locating}
Kevin Meng, David Bau, Alex Andonian, and Yonatan Belinkov.
\newblock Locating and editing factual associations in GPT.
\newblock \emph{Advances in Neural Information Processing Systems}, 2022.

\bibitem[Panickssery et~al.(2024)]{panickssery2024caa}
Nina Panickssery, Nick Gabrieli, Julian Schulz, Meg Tong, Evan Hubinger, and
Alexander Matt Turner.
\newblock Steering Llama 2 via contrastive activation addition.
\newblock \emph{arXiv preprint arXiv:2312.06681}, 2024.

\bibitem[Park et~al.(2024)]{park2024linear}
Kiho Park, Yo Joong Choe, and Victor Veitch.
\newblock The linear representation hypothesis and the geometry of large
language models.
\newblock In \emph{International Conference on Machine Learning}, 2024.

\bibitem[Perez et~al.(2023)]{perez2023sycophancy}
Ethan Perez et~al.
\newblock Simple synthetic data reduces sycophancy in large language models.
\newblock \emph{arXiv preprint arXiv:2308.03958}, 2023.

\bibitem[Qwen Team(2025)]{qwen2025technical}
Qwen Team.
\newblock Qwen3 technical report.
\newblock \emph{arXiv preprint arXiv:2505.09388}, 2025.

\bibitem[Sharma et~al.(2023)]{sharma2023sycophancy}
Mrinank Sharma, Meg Tong, Tomasz Korbak, David Duvenaud, Amanda Askell, Samuel
R. Bowman, and Ethan Perez.
\newblock Towards understanding sycophancy in language models.
\newblock \emph{arXiv preprint arXiv:2310.13548}, 2023.

\bibitem[Suzgun et~al.(2023)]{suzgun2023bbh}
Mirac Suzgun, Nathan Scales, Nathanael Sch\"arli, Sebastian Gehrmann, Yi Tay,
Hyung Won Chung, Aakanksha Chowdhery, Quoc V. Le, Ed H. Chi, Denny Zhou, and
Jason Wei.
\newblock Challenging BIG-Bench tasks and whether chain-of-thought can solve
them.
\newblock In \emph{Findings of ACL}, 2023.

\bibitem[Turner et~al.(2024)]{turner2024activation}
Alexander Matt Turner, Lisa Thiergart, Gavin Leech, David Udell, Juan J. Vazquez,
Ulisse Mini, and Monte MacDiarmid.
\newblock Activation addition: Steering language models without optimization.
\newblock \emph{arXiv preprint arXiv:2308.10248}, 2024.

\bibitem[Vennemeyer et~al.(2025)]{vennemeyer2025sycophancy}
Daniel Vennemeyer, Phan Anh Duong, Tiffany Zhan, and Tianyu Jiang.
\newblock Sycophancy is not one thing: Causal separation of sycophantic
behaviors in LLMs.
\newblock \emph{arXiv preprint arXiv:2509.21305}, 2025.

\bibitem[Wang et~al.(2023)]{wang2023interpretability}
Kevin Wang, Alexandre Variengien, Arthur Conmy, Buck Shlegeris, and Jacob
Steinhardt.
\newblock Interpretability in the wild: A circuit for indirect object
identification in GPT-2 small.
\newblock In \emph{International Conference on Learning Representations}, 2023.

\bibitem[Weng et~al.(2025)]{weng2025benchform}
Zhiyuan Weng, Guikun Chen, and Wenguan Wang.
\newblock Do as we do, not as you think: The conformity of large language
models.
\newblock In \emph{International Conference on Learning Representations}, 2025.

\bibitem[Zhong et~al.(2025)]{zhong2025disentangling}
Huixin Zhong, Yanan Liu, Qi Cao, Shijin Wang, Zijing Ye, Zimu Wang, and Shiyao
Zhang.
\newblock Disentangling the drivers of LLM social conformity: an
uncertainty-moderated dual-process mechanism.
\newblock \emph{arXiv preprint arXiv:2508.14918}, 2025.

\bibitem[Zhu et~al.(2025)]{zhu2025conformity}
Xiaochen Zhu, Caiqi Zhang, Tom Stafford, Nigel Collier, and Andreas Vlachos.
\newblock Conformity in large language models.
\newblock \emph{arXiv preprint arXiv:2410.12428}, 2025.

\bibitem[Zou et~al.(2023)]{zou2023representation}
Andy Zou, Long Phan, Sarah Chen, James Campbell, Phillip Guo, Richard Ren,
Alexander Pan, Xuwang Yin, Mantas Mazeika, Ann-Kathrin Dombrowski, Shashwat
Goel, Nathaniel Li, Michael J. Byun, Zifan Wang, Alex Mallen, Steven Basart,
Sanmi Koyejo, Dawn Song, Matt Fredrikson, J. Zico Kolter, and Dan Hendrycks.
\newblock Representation engineering: A top-down approach to AI transparency.
\newblock \emph{arXiv preprint arXiv:2310.01405}, 2023.

\end{thebibliography}
}

\appendix

\newpage
\input{checklist_filled.tex}

\end{document}
